---
title: "Using Realwage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using Realwage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=6,
  fig.height = 4
)
```

```{r setup}
library(realwage)
```

This short vignette aims to provide some inspiration around possible uses for the `realwage` package. I hope you find it useful.

## Getting Started

The simplest possible way to use `realwage` is to create a tibble of past salaries with *year*, *month*, and *salary* columns and then use that tibble to make a ggplot showing the salaries adjusted for inflation. This can be done with the following two commands. It assumes both the `realwage` and `tibble` packages are installed and that the `realwage` package is loaded.

```{r}
wage_table <- tibble::tibble(year = c(2020, 2021, 2022),
                             month = c(3, 3, 3),
                             salary = c(30000, 30250, 31000))
print(wage_table)
```

```{r}
calculate_and_plot_adjusted_salary(wage_table)
```

## Changing Date Range

If you only want to plot values for a particular range of dates `calculate_and_plot_adjusted_salary()` accepts `start_date` and `end_date` parameters in ISO format (i.e. yyyy-mm-dd).

```{r}
calculate_and_plot_adjusted_salary(wage_table, end_date = '2021-10-01')
```

## Making Other Changes to Graph

The ggplot object returned by `calculate_and_plot_adjusted_salary()` can be saved and modified as any ggplot.

```{r}
wage_plot <- calculate_and_plot_adjusted_salary(wage_table)

wage_plot +
  ggplot2::ggtitle('My Salary Over the Years')
```

## Working Directly with the Underlying Data

To download and work directly with the data, before plotting it, use `make_inflation_adjusted_wage_table()`. The data is in tidy format and can then be processed with all the usual tidyverse tools.

```{r}
inflation_data <- make_inflation_adjusted_wage_table(wage_table)
print(inflation_data)
```

If you, for example, want to remove all but one inflation calculation, this can be accomplished using `dplyr`.

```{r}
inflation_data <- dplyr::filter(inflation_data,
                                series %in% c('Starting March 2020', 'Actual Salary'))
```

This data can of course be used however you want, e.g. to make your own plots using `ggplot2`, but if you want to make a plot similar to the ones above you can use `plot_adjusted_wages()`.

```{r}
plot_adjusted_wages(inflation_data)
```

To make the data more human readable, it can be pivoted using `tidyr`.

```{r}
tidyr::pivot_wider(inflation_data, names_from = series)
```

## Downloading Consumer Price Index

You can use `get_cpi()` to download [Consumer Price Index data from Statistics Sweden](http://www.statistikdatabasen.scb.se/pxweb/sv/ssd/START__PR__PR0101__PR0101A/KPIFastM2) to use for your own calculations.

```{r}
get_cpi()
```

Note that they have many other datasets, including CPI data going back to the 19th century. The dataset used in the `realwage` package only goes back to 2006, but was chosen as it provides consistently formatted monthly figures.
